### 2018-09-09 paper/finitecomplete-failures.tex
Technical report completed, taking into account the remarks from Ana

% --------------------------------------------------------------------------
\subsubsection*{Explanation of the Test Definition}
A test is performed by running $U_F(p)$ concurrently with any SUT process
$Q$, synchronising over $\Sigma$. So, a test execution is a trace of
the concurrent process
\[
Q\parallel[\Sigma] U_F(p).
\]
It is assumed that the events $\efail$ and $\epass$, indicating a verdict
FAIL and PASS for the test execution, are not included in $\Sigma$. Since we
assume that $Q$ is free of livelocks, it is guaranteed that events 
$\efail$ or $\epass$ always become visible, if they are the only events
$U_F(p)/s$ is ready to engage in.

The test is
\emph{passed} by the SUT (written $Q\ \pass\ U_F(p)$) if, and only if, {\it
every} execution of $Q\parallel[\Sigma] U_F(p)$ terminates with   event
$\epass$. This can also be  expressed by means of a failures refinement as
defined below.
%
\begin{equation}
\label{eq:passF}
Q\ \pass\ U_F(p) \defs (\epass\then\Stop) \lessdet_F (Q\parallel[\Sigma] U_F(p)) \hide \Sigma
\end{equation}
%
This type of pass relation is often called \emph{must test}, because every
test execution must end with the $\epass$
event~\cite{Hennessy:1988:ATP:50497}. Note that it is necessary to use the
failures-refinement relation in this condition, and not the trace-refinement
relation:~$(Q\parallel[\Sigma] U_F(p)) \hide \Sigma$ may have  the same
visible traces $\varepsilon$ and $\epass$ as the ``Test Passed
Process'' $(\epass\then\Stop)$. However, the former may nondeterministically
refuse $\epass$, due to a deadlock occurring when a faulty SUT process
executes concurrently with $U_F(p,k,n)$ executing branch (\ref{eq:ufd}),
because $(k = p\wedge \minhits(P/s)\neq\varnothing)$. 
This is explained further in the next paragraphs.

The specification of $U_F(p,k,n)$ implies that the test always stops after having engaged into a trace $s\in\trc(Q)$ of maximal length $p$ or $p+1$: 
If branch (\ref{eq:ufa})
is the last to be entered, the maximal length of $s$ is $p+1$, and the test execution
stops with $\efail$. If branch (\ref{eq:ufb}) is the last to be entered, the maximal length of $s$ is $p$, and the execution stops with $\epass$. If branch (\ref{eq:ufd})
is the last to be entered, the process either accepts another event
$e$ of some minimal hitting set $H\in\minhits(P/front(s))$. Then the final length
of $s$ is $p+1$, and the execution terminates with $\epass$. Or the test execution
$(Q\parallel[\Sigma] U_F(p))/s$ deadlocks,  the final length of $s$ is $p$,
and the execution stops without a PASS or FAIL event. Such an execution is also interpreted as FAIL, because it reveals that $(\epass\then\Stop) \not\lessdet_F (Q\parallel[\Sigma] U_F(p)) \hide \Sigma$.

We observe that the number of possible executions of $Q\parallel[\Sigma]
U_F(p)$ is finite, because the number of traces $s$ with maximal length
$(p+1)$ is finite and the sets $[n]^0$, $(\Sigma - [n]^0)$, and
$\minhits(n)$ are finite. We further recall that $\minhits(n)$ may be
empty, so that the indexed internal choice construct in (\ref{eq:ufd}) 
is undefined. Therefore, the associated guard condition contains the requirement
$\minhits(P/s)\neq\varnothing$, otherwise this branch cannot be taken,
but branches (\ref{eq:ufa}) or (\ref{eq:ufb}) can be taken in this situation.
%%%\fixme{alcc: Oops. This is undefined in CSP.}

Intuitively speaking, $U_F(p)$ is able to perform any trace $s$ of $P$, up to
a length $p$. If, after having already run through $s\in\trc(P)$ with
$\#s \le p$, an event is accepted by the SUT that is outside the initials of $P/s$
 (recall from Lemma~\ref{lemma:ufproperties} that $[n]^0 = [P/s]^0$ for $U_F(p)/s$),
the test immediately terminates with FAIL-event $\efail$. This is handled by
the branch (\ref{eq:ufa}) of the external choice in the process $U_F(p,s)$
defined above.

If $P/s$ is the $\Stop$ process or has $\Stop$ as an internal choice,
this is revealed by $\minhits(n) = \varnothing$ (recall (\ref{eq:minhitminaccempty}) and
Lemma~\ref{lemma:ufproperties}). 
In this case, the test may terminate successfully (branch (\ref{eq:ufb}) of
the external choice in $U_F(p,s)$). If $P/s$ may also nondeterministically engage into events in such a situation, branches  (\ref{eq:ufa}) and (\ref{eq:ufc}) of the test are
simultaneously enabled.

If the length of $s$ is still less than $p$, the test accepts any event $e$ from
the initials $[P/s]^0 = [G(P)/s]^0$ and continues recursively as $U_F(p,\#s+1,G(P)/s.e)$ in
branch~(\ref{eq:ufc}), this follows again from  Lemma~\ref{lemma:ufproperties}). 
A test of this type is called \emph{adaptive}, because
it accepts any legal behaviour of the SUT and adapts its consecutive
behaviour to the event selected by the SUT.

After having run through a trace of length $p$, and if
$\minhits(P/s)\neq \varnothing$, the test changes its behaviour: instead of
offering {\it all} legal events from $[P/s]^0$ to the SUT, it
nondeterministically chooses a minimal hitting set   $H\in \minhits(G(P)/s)$ 
and
only offers the events contained in $H$. If the SUT refuses to engage
into any event of $H$, this reveals a violation of failures refinement:
according to Lemma~\ref{lemma:hseta}, a conforming SUT should accept at least
one event of each minimal hitting set in $\minhits(P/s)$. Therefore, the test
only terminates with  $\epass$, if such an event is accepted by the SUT.

@@

\begin{theorem}\label{th:failurestest}
Let $P$ be a divergence-free CSP process over alphabet $\Sigma$ whose
normalised transition graph $G(P)$ has $p$ states. Define fault domain ${\cal
D}$ as the set of all divergence-free CSP processes over alphabet $\Sigma$,
whose transition graph has at most $q$ states with $q \ge p$. Then the test
suite
\[
\TS_F = \{ U_F(k)~|~0 \le k < pq  \}\quad\text{$U_F(k)$ specified in (\ref{eq:UFP})}
\]
is complete with respect to ${\cal F} = (P,\lessdet_F,{\cal D})$.
\end{theorem}
%
The proof of the theorem follows from the two lemmas below. The first states
that test suite $\TS_F$ is sound, the second states that the suite is also
exhaustive.
%
\begin{lemma}\label{lemma:mainfsound}
Test suite $\TS_F$ generated from a CSP process $P$, as specified in
Theorem~\ref{th:failurestest}, is passed by every CSP process $Q$ satisfying
$P\lessdet_F Q$.
\end{lemma}
\begin{proof}{~}\newline
{\bf Step~1.} Suppose that $P\lessdet_F Q$, so $P\lessdet_T Q$ and  $Q\ conf\
P$ according to (\ref{eq:failconf}). Since   $\trc(Q)\subseteq \trc(P)$,
any adaptive test $U_F(p)$ running in parallel with $Q$ will always enter the
branches (\ref{eq:ufb}), (\ref{eq:ufc}), or (\ref{eq:ufd}) of the external
choice construction for $U_F(p,k,n)$:  
If $U_F(p,k,n) = U_F(p)/s$ with $s\in\trc(P)$, then Lemma~\ref{lemma:ufproperties}
implies $U_F(p,k,n) = U_F(p,k,G(P)/s)$, so $[n]^0 = [G(P)/s]^0 = [P/s]^0$. As
a consequence, $[Q/s]^0\subseteq [P/s]^0 = [n]^0$, so branch~(\ref{eq:ufa}) can never be
entered in the parallel execution of $Q$ and $U_F(p)$.

\medskip
\noindent
{\bf Step~2.} Sinc $Q\ conf\ P$, Lemma~\ref{lemma:hseta} implies that for all  
$s\in\trc(Q)\cap\trc(P)$, every $H \in \minhits(P/s)$ is a hitting set for
$\minaccs(Q/s)$. 
From Lemma~\ref{lemma:ufproperties} we know that
$U_F(p)/s = U_F(p,k,n)$ with $U_F(p,k,n) = U_F(p,\#s,G(P)/s)$, 
so $\minhits(n) = \minhits(P/s)$. 
Branch (\ref{eq:ufb}) of test $U_F(p,k,n)$ leads always to a
PASS verdict and is taken if $\minhits(P/s) = \varnothing$. Branch (\ref{eq:ufc}) 
always leads to test continuation without a
verdict. For the last branch, we note that it can only be entered if $\minhits(P/s)
= \minhits(n) \neq\varnothing$. In this case,
any selected minimal hitting set
$H\in\minhits(n)$ has a non-empty intersection with each of the minimal
acceptances of $Q/s$. 
As a consequence, $Q/s$ never blocks when offered
events from $H\in\minhits(n)$, 
and the test terminates with PASS event $\epass$. Note that
this argument requires that $Q$ is free of livelocks, because otherwise the
$\epass$-events might not become visible, due to unbounded sequences of
hidden events performed by $Q$. 
\xbox
\end{proof}
%
\begin{lemma}\label{lemma:mainfexhaustive}
Test suite $\TS_F$ specified in Theorem~\ref{th:failurestest} is exhaustive
for the fault model specified there.
\end{lemma}
\begin{proof}
Consider a process $Q\in{\cal D}$ with $P\not\lessdet_F Q$, According to
(\ref{eq:failconf}), this non-conformance can be caused in two possible ways
corresponding to the cases $P\not\lessdet_T Q$ and $\neg(Q\ conf\ P)$,
respectively:
\begin{description}
\item[Case~1] $\trc(Q)\not\subseteq \trc(P)$
\item[Case~2] There exists a joint trace $s\in\trc(Q)\cap\trc(P)$ and a minimal acceptance $A_Q$
of $\minaccs(Q/s)$, such that
(see Lemma~\ref{lemma:tgtrcref}, (\ref{eq:failrefb})).
\begin{equation}
\label{eq:accsnotcontained}
\forall A_P\in\minaccs(P/s): A_P\not\subseteq A_Q,
\end{equation}
\end{description}
It has to be shown for each of the two cases that at least one test
execution of some $(Q\parallel[\Sigma] U_F(k))$ with $k < pq$ ends with the
   $\efail$-event or without giving any verdict. The latter case is also
interpreted  as FAIL, since then the process $\epass\then\Stop$ is no longer
failures-refined by $(Q\parallel[\Sigma] U_F(k))\hide \Sigma$.

\medskip
\noindent
{\bf Case~1.} Consider a  trace $s.e\in\trc(Q)$ with $s\in\trc(P)$, but
$s.e\not\in\trc(P)$. Such a trace always exists because $\varepsilon$ is a
trace of every process. In this case, $s$ is also a trace of the product
graph $G = G(P)\times G(Q)$ defined in Section~\ref{sec:ntg}. Suppose that
$G/s = (n_P,n_Q)$. The length of $s$ is not known, but from the construction
of $G$,  we know that $G$ has at most $pq$ reachable states, because $G(P)$
has $p$ states, and $G(Q)$ has at most $q$ states. By
Lemma~\ref{lemma:reachproduc}, $(n_P,n_Q)$ can be reached by a trace
$u\in\trc(G)$ of length $\#u < pq$. Now the construction of the transition
function of $G$ implies that $u$ is also a trace of $P$ and $Q$. Since test
$U_F(pq-1)$ accepts all traces of $P$ up to length $pq-1$, $u$ is also a
trace of this test, and, by construction and by Lemma~\ref{lemma:ufproperties}, 
$U_F(pq-1)/u = U_F(pq-1,\#u,G(P)/u)$. Since
$s.e\not\in\trc(P)$, $e$ is an element of $\Sigma-[P/u]^0 = \Sigma - [G(P)/s]^0$. 
Hence, in at least
one execution, $U_F(pq-1,\#u,G(P)/u)$ executes its first branch (\ref{eq:ufa}) with
this event $e$, so that the test fails. Again, the
assumption of non-divergence of Q is needed for this conclusion.

\medskip
\noindent
{\bf Case~2.} We note that trace $s$ is again a trace of the product graph
$G$, but we do not know its length. Again, by applying Lemma~\ref{lemma:reachproduc},
we know that the state   $G/s$ can   be reached by a trace
$u\in\trc(Q)\cap\trc(P)$ of maximal length $\#u < pq$. Consider test $U_F(\#
u)$, which satisfies $U_F(\# u)/u = U_F(\#u,\#u,G(P)/u)$, because of Lemma~\ref{lemma:ufproperties}. This process   always performs
branch (\ref{eq:ufc}) until the trace $u$ has been completely processed.
$U_F(\#u,\#u,G(P)/u)$ may execute branches (\ref{eq:ufa}) or (\ref{eq:ufd})
only:~assumption (\ref{eq:accsnotcontained}) in Case~2 implies that $P/s$ has
at least one non-empty minimal acceptance. From (\ref{eq:minhitminaccempty}) 
we know that this is equivalent to $\minhits(P/s) = \minhits(G(P)/s)\neq\varnothing$.
As a consequence, branch (\ref{eq:ufb}) can never be taken because its guard condition
evaluates to $\isf$  for $U_F(\#u,\#u,G(P)/u)$.
Moreover, the guard condition $(k < p)$ for branch (\ref{eq:ufc}) evaluates
to $\isf$ for $U_F(\#u,\#u,G(P)/u)$, too. If branch (\ref{eq:ufa}) is executed, the
test always fails. If branch (\ref{eq:ufd}) is executed, the test fails for
the execution where a minimal hitting set $H\in\minhits(P/u)$ is chosen
by $U_F(\#u,\#u,G(P)/u)$ that has an empty intersection with the minimal acceptance
$A_Q$ from condition (\ref{eq:accsnotcontained}). The existence of such an
$H$ is guaranteed because of Lemma~\ref{lemma:hseta}. As a consequence, there
exists a test execution   where $Q/u$ selects acceptance $A_Q$ and
$U_F(\#u,\#u,G(P)/u)$ selects $H$. This execution deadlocks in process state
$(Q\parallel[\Sigma]U_F(\# u))/u$, so it cannot produce the $\epass$-event;
this  means that the test fails and concludes the proof. 
\xbox
\end{proof}

% ==========================================================================
\subsection{Complexity Considerations}
\label{sec:complexity}

As can be seen from the specification of the test cases $U_F(k)$, the number of
executions ending in a $\epass$-event corresponds to the number $\ell$ of traces $s$
of $P$ with length equal to\footnote{In this estimation, we disregard
the case where the test
terminates earlier due to entering branch (\ref{eq:ufb}).} $k$,
multiplied by the number $h$ of minimal hitting sets in
$\minhits(P/s)$.

% -------------------------------------------------------------------------
\subsubsection*{Estimation of $\ell$}
The first factor $\ell$
has worst-case upper bound $\ell\le |\Sigma|^k$. As an example, where this upper bound 
is really met, consider the reference process
\[
RUN(\Sigma) = e:\Sigma \then RUN(\Sigma)
\]
The normalised transition graph of this process has a single state, and its initials
are $[RUN(\Sigma)]^0 = \Sigma$. Therefore, the associated test process $U_F(k)$ can 
never enter branches (\ref{eq:ufa}) and (\ref{eq:ufb}), but there are exactly 
$|\Sigma|^k$ different traces of length $k$ exercising branch (\ref{eq:ufc})
for each of their events.

% -------------------------------------------------------------------------
\subsubsection*{Estimation of $h$}
Given a set $\minaccs(P/s) = \{ A_1,\dots, A_\alpha \}$ of   minimal
acceptances, the cardinality $h$ of $\minhits(P/s)$ is maximal for the case
where all $A_i$ are disjoint. In this case,  $h~~=~~|\minhits(P/s)|~~=~~
\prod_{i=1}^\alpha|A_i|$. To find an upper bound for $h$, we use the result
that we prove in the following lemma.
%
\begin{lemma}\label{lemma:minhitmax}
Let $n\ge 2$ be any positive integer. We call a sequence
%
$$0 < a_1\le a_2\le \ldots\le  a_{\alpha}, a_i\in \mathbb{N}, i=1,\dots, \alpha$$
%
a partition of $n$, if $n=\Sigma_{i=1}^{\alpha}a_i$. Define
%
$$p(n)=\begin{cases} 3^k &\text{if}\,\, n=3k\\ 2^2\cdot 3^{k-1} &\text{if}\,\, n=3k+1\\2\cdot 3^{k} &\text{if}\,\, n=3k+2\end{cases}$$
Then  $$p(n)=\max\big\{\prod_{i=1}^{\alpha}a_i~|~n=\Sigma_{i=1}^{\alpha}a_i,\ a_i,
\alpha\in \mathbb{N}\big\}.$$
%
\end{lemma}
\begin{proof}
Let $n\ge 2$. Obviously, \begin{equation}\label{eq:1} n\le
p(n).\end{equation} Define $\pi(n)$ to be the following partition  of $n$:
%
$$\pi(n)=\begin{cases} \underbrace{3,\ldots,3}_{k\text{ times}}  &\text{if}\,\, n=3k\\ 2,2,\underbrace{3,\ldots,3}_{k-1\text{ times}}  &\text{if}\,\, n=3k+1\\2,\underbrace{3,\ldots,3}_{k\text{ times}}  &\text{if}\,\, n=3k+2\end{cases}$$
%
Then the product of the numbers in $\pi(n)$ equals  $p(n)$. In the remainder
of the proof we have to show that $p(n)$ is really maximal in the sense of
the lemma.

Since the number of partitions of a given $n$ is finite,
%
$$\text{max}\{\prod_{i=1}^{\alpha}a_i~|~n=\Sigma_{i=1}^{\alpha}a_i, a_i\in \mathbb{N}\}$$
%
exists. Let $a_1, \ldots , a_{\alpha}$ be any partition of $n\ge 2$ with
maximal product. Then, obviously,  $a_i>1$ for all $i=1,\dots,\alpha$.
Suppose there is some $a_j$, $1\le j\le \alpha$, with $a_j>3$. Then
from~(\ref{eq:1}) we have
$$\prod_{i=1}^{\alpha}a_i\le a_1\cdot a_{j-1}\cdot p(a_j)\cdot a_{j+1}\ldots\cdot a_{\alpha}.$$
Hence we can replace $a_j$ by $\pi(a_j)$, and the new partition has  a
product value which is not less then the original partition $a_1, \ldots ,
a_{\alpha}$. This process can be repeated, until every $a_i$ in the resulting
partition is either 2 or 3. Hence there exists a partition of $n$ with
maximal product, such that every term is $2$ or $3$. Since $6=2+2+2=3 + 3$
and $2^3< 3^2$, the number of $2$'s in such a maximal partition is less than
$3$. Let $a_1\le a_2\le \ldots\le  a_{\alpha}$ with $a_i\in \{2,3\},
i=1,\dots,\alpha$, be a partition of $n$ with maximal product. Let $k_1\le 2$
be the number of terms equal $2$ and $k_2$ be the number of terms equal $3$.
Every $n\ge 2$ can be represented by $n=2k_1+3k_2$, $n\equiv 2k_1 (\mod 3)$
with $k_1\le 2$, and $\prod_{i=1}^{\alpha}a_i=2^{k_1}\cdot 3^{k_2}$. We have

\begin{align*}k_1&=\begin{cases} 0 &\text{if}\,\, n=3k\\ 2 &\text{if}\,\, n=3k+1\\1 &\text{if}\,\, n=3k+2\end{cases}\\
 k_2&=\begin{cases} k &\text{if}\,\, n=3k\\ k-1 &\text{if}\,\, n=3k+1\\k &\text{if}\,\, n=3k+2\end{cases}\\
\prod_{i=1}^{\alpha}a_i&=\begin{cases} 3^k &\text{if}\,\, n=3k\\ 2^2\cdot 3^{k-1} &\text{if}\,\, n=3k+1\\2\cdot 3^{k} &\text{if}\,\, n=3k+2\end{cases}\end{align*}
This proves the lemma.
\xbox
\end{proof}
%
From Lemma~\ref{lemma:minhitmax}, we conclude that
$$h=\begin{cases}
3^{\lfloor\frac{|\Sigma|}{3}\rfloor} &\text{if}\  |\Sigma|~~\equiv~~0 \mod 3\\
3^{\lfloor\frac{|\Sigma|}{3}\rfloor}  + 3^{\lfloor\frac{|\Sigma|}{3}\rfloor-1}&\text{if}\ |\Sigma|~~\equiv~~1\mod 3\\
2\cdot 3^{\lfloor\frac{|\Sigma|}{3}\rfloor} &\text{if}\ |\Sigma|~~\equiv~~2 \mod 3
\end{cases}
$$
%
We note that $h$ is significantly smaller than $2^{|\Sigma|}-1$, this can
also be intuitively motivated by the fact that  for every $H\in
\minhits(P/s)$, all of its true subsets and true supersets are {\it not}
contained in $\minhits(P/s)$.

>>>>

% --------------------------------------------------------------------------
\subsubsection*{Explanation of the Test Definition}
A test is performed by running $U_F(p)$ concurrently with any SUT process
$Q$, synchronising over $\Sigma$. So, a test execution is a trace of
the concurrent process
\[
Q\parallel[\Sigma] U_F(p).
\]
It is assumed that the events $\efail$ and $\epass$, indicating a verdict
FAIL and PASS for the test execution, are not included in $\Sigma$. Since we
assume that $Q$ is free of livelocks, it is guaranteed that events 
$\efail$ or $\epass$ always become visible, if they are the only events
$U_F(p)/s$ is ready to engage in: if $U_F(p)/s$ can only produce $\epass$ or
$\efail$, the occurrence of these events can never be blocked due to a livelock in $Q$
occurring in the same step of the execution.

The test is
\emph{passed} by the SUT (written $Q\ \pass\ U_F(p)$) if, and only if, {\it
every} execution of $Q\parallel[\Sigma] U_F(p)$ terminates with   event
$\epass$. This can also be  expressed by means of a failures refinement as
defined below.
%
\begin{equation}
\label{eq:passF}
Q\ \pass\ U_F(p) \defs (\epass\then\Stop) \lessdet_F (Q\parallel[\Sigma] U_F(p)) \hide \Sigma
\end{equation}
%
This type of pass relation is often called \emph{must test}, because every
test execution must end with the $\epass$
event~\cite{Hennessy:1988:ATP:50497}. Note that it is necessary to use the
failures-refinement relation in this condition, and not the trace-refinement
relation:~$(Q\parallel[\Sigma] U_F(p)) \hide \Sigma$ may have  the same
visible traces $\varepsilon$ and $\epass$ as the ``Test Passed
Process'' $(\epass\then\Stop)$. However, the former may nondeterministically
refuse $\epass$, due to a deadlock occurring when a faulty SUT process
executes concurrently with $U_F(p,k,n)$ executing branch (\ref{eq:ufd}),
when guard condition $(k = p\wedge \minhits(n)\neq\varnothing)$ evaluates to $\ist$. 
This is explained further in the next paragraphs.

The specification of $U_F(p,k,n)$ implies that the test always stops after having engaged into a trace $s\in\trc(Q)$ of maximal length $p$ or $p+1$: 
If branch (\ref{eq:ufa})
is the last to be entered, the maximal length of $s$ is $p+1$, and the test execution
stops with $\efail$. If branch (\ref{eq:ufb}) is the last to be entered, the maximal length of $s$ is $p$, and the execution stops with $\epass$. If branch (\ref{eq:ufd})
is the last to be entered, the process either accepts another event
$e$ of some minimal hitting set $H\in\minhits(n)$ with $n = G(P)/s$ according to
Lemma~\ref{lemma:ufproperties}. Then the final length
of $s$ is $p+1$, and the execution terminates with $\epass$. Or the test execution
$(Q\parallel[\Sigma] U_F(p))/s$ deadlocks,  the final length of $s$ is $p$,
and the execution stops without a PASS or FAIL event. Such an execution is also interpreted as FAIL, because it reveals that $(\epass\then\Stop) \not\lessdet_F (Q\parallel[\Sigma] U_F(p)) \hide \Sigma$.

We observe that the number of possible executions of $Q\parallel[\Sigma]
U_F(p)$ is finite, because the number of traces $s$ with maximal length
$(p+1)$ is finite and the sets $[n]^0$, $(\Sigma - [n]^0)$, and
$\minhits(n)$ are finite. We further recall that $\minhits(n)$ may be
empty, so that the indexed internal choice construct in (\ref{eq:ufd}) 
is undefined. Therefore, the associated guard condition contains the requirement
$\minhits(n)\neq\varnothing$, otherwise this branch cannot be taken,
but branches (\ref{eq:ufa}) or (\ref{eq:ufb}) can be taken in this situation.
%%%\fixme{alcc: Oops. This is undefined in CSP.}

Intuitively speaking, $U_F(p)$ is able to perform any trace $s$ of $P$, up to
a length $p$. If, after having already run through $s\in\trc(P)$ with
$\#s \le p$, an event is accepted by the SUT that is outside the initials of $P/s$
 (recall from Lemma~\ref{lemma:ufproperties} that $[n]^0 = [P/s]^0$ for $U_F(p)/s$),
the test immediately terminates with FAIL-event $\efail$. This is handled by
the branch (\ref{eq:ufa}) of the external choice in the process $U_F(p,\#s,n)$
defined above.

If $P/s$ is the $\Stop$ process or has $\Stop$ as an internal choice,
this is revealed by $\minhits(G(P)/s) = \varnothing$ (recall (\ref{eq:minhitminaccempty}) and
Lemma~\ref{lemma:ufproperties}). 
In this case, the test may terminate successfully (branch (\ref{eq:ufb}) of
the external choice in $U_F(p,\#s,G(P)/s)$). 
If $P/s$ may also nondeterministically engage into events in such a situation, branche 
 (\ref{eq:ufc}) of the test is
simultaneously enabled. If $Q/s$ is able to engage into an event from 
$\Sigma - [P/s]^0$, a test execution exists where $U_F(p,\#s,G(P)/s)$ branches
into (\ref{eq:ifpb}) and produces the $\efail$-event.

If the length of $s$ is still less than $p$, the test accepts any event $e$ from
the initials $[P/s]^0 = [G(P)/s]^0$ and continues recursively as $U_F(p,\#s+1,G(P)/s.e)$ in
branch~(\ref{eq:ufc}), this follows again from  Lemma~\ref{lemma:ufproperties}. 
A test of this type is called \emph{adaptive}, because
it accepts any legal behaviour of the SUT and adapts its consecutive
behaviour to the event selected by the SUT.

Suppose that a test execution has run through a trace 
$s\in\trc(P)$ of length $p$, so that $U_F(p)/s = U_F(p,p,n)$ with $n = G(P)/s$. 
Then,   if
$\minhits(n)\neq \varnothing$, the test changes its behaviour: instead of
offering {\it all} legal events from $[n]^0$ to the SUT, it
nondeterministically chooses a minimal hitting set   $H\in \minhits(n)$ 
and
only offers the events contained in $H$. If the SUT refuses to engage
into any event of $H$, this reveals a violation of failures refinement:
according to Lemma~\ref{lemma:hseta}, a conforming SUT should accept at least
one event of each minimal hitting set in $\minhits(n)$. Therefore, the test
execution
only terminates with  $\epass$, if such an event is accepted by the SUT.

@@

\begin{theorem}\label{th:failurestest}
Let $P$ be a non-terminating, divergence-free CSP process over alphabet $\Sigma$ whose
normalised transition graph $G(P)$ has $p$ states. Define fault domain ${\cal
D}$ as the set of all divergence-free CSP processes over alphabet $\Sigma$,
whose transition graph has at most $q$ states with $q \ge p$. Then the test
suite
\[
\TS_F = \{ U_F(k)~|~0 \le k < pq  \}\quad\text{$U_F(k)$ specified in (\ref{eq:UFP})}
\]
is complete with respect to ${\cal F} = (P,\lessdet_F,{\cal D})$.
\end{theorem}
%
The proof of the theorem follows from the two lemmas below. The first states
that test suite $\TS_F$ is sound, the second states that the suite is also
exhaustive.
%
\begin{lemma}\label{lemma:mainfsound}
Test suite $\TS_F$ generated from a CSP process $P$, as specified in
Theorem~\ref{th:failurestest}, is passed by every CSP process $Q$ satisfying
$P\lessdet_F Q$.
\end{lemma}
\begin{proof}{~}\newline
{\bf Step~1.} Suppose that $P\lessdet_F Q$, so $P\lessdet_T Q$ and  $Q\ conf\
P$ according to (\ref{eq:failconf}). Since   $\trc(Q)\subseteq \trc(P)$,
any adaptive test $U_F(p)$ running in parallel with $Q$ will always enter the
branches (\ref{eq:ufb}), (\ref{eq:ufc}), or (\ref{eq:ufd}) of the external
choice construction for $U_F(p,k,n)$:  
If $U_F(p,k,n) = U_F(p)/s$ with $s\in\trc(P)$, then Lemma~\ref{lemma:ufproperties}
implies $U_F(p,k,n) = U_F(p,k,G(P)/s)$, so $[n]^0 = [G(P)/s]^0 = [P/s]^0$. As
a consequence, $[Q/s]^0\subseteq [P/s]^0 = [n]^0$, so branch~(\ref{eq:ufa}) can never be
entered in the parallel execution of $Q$ and $U_F(p)$.

\medskip
\noindent
{\bf Step~2.} Sinc $Q\ conf\ P$, Lemma~\ref{lemma:hseta} implies that for all  
$s\in\trc(Q)\cap\trc(P)$, every $H \in \minhits(P/s)$ is a hitting set for
$\minaccs(Q/s)$. 
From Lemma~\ref{lemma:ufproperties} we know that
$U_F(p)/s = U_F(p,k,n)$ with $U_F(p,k,n) = U_F(p,\#s,G(P)/s)$, 
so $\minhits(n) = \minhits(P/s)$. 
Branch (\ref{eq:ufb}) of test $U_F(p,k,n)$ leads always to a
PASS verdict and is taken if $\minhits(n) = \varnothing$. Branch (\ref{eq:ufc}) 
always leads to test continuation without a
verdict. For the last branch, we note that it can only be entered if $\minhits(P/s)
= \minhits(n) \neq\varnothing$. In this case,
any selected minimal hitting set
$H\in\minhits(n)$ has a non-empty intersection with each of the minimal
acceptances of $Q/s$. 
As a consequence, $Q/s$ never blocks when offered
events from $H\in\minhits(n)$, 
and the test terminates with PASS event $\epass$. Note that
this argument requires that $Q$ is free of livelocks, because otherwise the
$\epass$-events might not become visible, due to unbounded sequences of
hidden events performed by $Q$. 
\xbox
\end{proof}
%
\begin{lemma}\label{lemma:mainfexhaustive}
Test suite $\TS_F$ specified in Theorem~\ref{th:failurestest} is exhaustive
for the fault model specified there.
\end{lemma}
\begin{proof}
Consider a process $Q\in{\cal D}$ with $P\not\lessdet_F Q$, According to
(\ref{eq:failconf}), this non-conformance can be caused in two possible ways
corresponding to the cases $P\not\lessdet_T Q$ and $\neg(Q\ conf\ P)$,
respectively:
\begin{description}
\item[Case~1] $\trc(Q)\not\subseteq \trc(P)$
\item[Case~2] There exists a joint trace $s\in\trc(Q)\cap\trc(P)$ and a minimal acceptance $A_Q$
of $\minaccs(Q/s)$, such that
(see Lemma~\ref{lemma:tgtrcref}, (\ref{eq:failrefb})).
\begin{equation}
\label{eq:accsnotcontained}
\forall A_P\in\minaccs(P/s): A_P\not\subseteq A_Q,
\end{equation}
\end{description}
It has to be shown for each of the two cases that at least one test
execution of some $(Q\parallel[\Sigma] U_F(k))$ with $k < pq$ ends with the
   $\efail$-event or without giving any verdict. The latter case is also
interpreted  as FAIL, since then the process $\epass\then\Stop$ is no longer
failures-refined by $(Q\parallel[\Sigma] U_F(k))\hide \Sigma$.

\medskip
\noindent
{\bf Case~1.} Consider a  trace $s.e\in\trc(Q)$ with $s\in\trc(P)$, but
$s.e\not\in\trc(P)$. Such a trace always exists because $\varepsilon$ is a
trace of every process. In this case, $s$ is also a trace of the product
graph $G = G(P)\times G(Q)$ defined in Section~\ref{sec:ntg}, and  
$G/s = (G(P)/s,G(Q)/s)$ holds. The length of $s$ is not known, but from the construction
of $G$,  we know that $G$ has at most $pq$ reachable states, because $G(P)$
has $p$ states, and $G(Q)$ has at most $q$ states. By
Lemma~\ref{lemma:reachproduc}, $(G(P)/s,G(Q)/s)$ can be reached by a trace
$u\in\trc(G)$ of length $\#u < pq$. Now the construction of the transition
function of $G$ implies that $u$ is also a trace of $P$ and $Q$, 
which means that 
$(G(P)/s,G(Q)/s) = (G(P)/u,G(Q)/u)$. Since test
$U_F(pq-1)$ accepts all traces of $P$ up to length $pq-1$, $u$ is also a
trace of this test, and, by construction and by Lemma~\ref{lemma:ufproperties}, 
$U_F(pq-1)/u = U_F(pq-1,\#u,G(P)/u)$. Since
$s.e\not\in\trc(P)$, $e$ is an element of $\Sigma-[P/u]^0 = \Sigma - [G(P)/s]^0$. 
Hence, in at least
one execution, $U_F(pq-1,\#u,G(P)/u)$ executes its first branch (\ref{eq:ufa}) with
this event $e$, so that the test fails. Again, the
assumption of non-divergence of Q is needed for this conclusion.

\medskip
\noindent
{\bf Case~2.} We note that trace $s$ is again a trace of the product graph
$G$, but we do not know its length. Again, by applying Lemma~\ref{lemma:reachproduc},
we know that the state   $G/s$ can   be reached by a trace
$u\in\trc(Q)\cap\trc(P)$ of maximal length $\#u < pq$. Consider test $U_F(\#
u)$, which satisfies $U_F(\# u)/u = U_F(\#u,\#u,G(P)/u)$, because of Lemma~\ref{lemma:ufproperties}. Test $U_F(\#u)$   always performs
branch (\ref{eq:ufc}) until the trace $u$ has been completely processed.
$U_F(\#u,\#u,G(P)/u)$ may execute branches (\ref{eq:ufa}) or (\ref{eq:ufd})
only:~assumption (\ref{eq:accsnotcontained}) in Case~2 implies that $P/s$ has
at least one non-empty minimal acceptance. From (\ref{eq:minhitminaccempty}) 
we know that this is equivalent to $\minhits(P/s) = \minhits(G(P)/s)\neq\varnothing$, and
we observe that $G(P)/s = G(P)/u$, so $\minhits(G(P)/u)\neq\varnothing$.
As a consequence, branch (\ref{eq:ufb}) cannot be taken because its guard condition
evaluates to $\isf$  for $U_F(\#u,\#u,G(P)/u)$.
The guard condition $(k < p)$ for branch (\ref{eq:ufc}) evaluates
to $\isf$ for $U_F(\#u,\#u,G(P)/u)$, too. If branch (\ref{eq:ufa}) is executed, the
test always fails. If branch (\ref{eq:ufd}) is executed, the test fails for
the execution where a minimal hitting set $H\in\minhits(P/u)$ is chosen
by $U_F(\#u,\#u,G(P)/u)$ that has an empty intersection with the minimal acceptance
$A_Q$ from condition (\ref{eq:accsnotcontained}). The existence of such an
$H$ is guaranteed because of Lemma~\ref{lemma:hseta}. As a consequence, there
exists a test execution   where $Q/u$ selects acceptance $A_Q$ and
$U_F(\#u,\#u,G(P)/u)$ selects $H$. This execution deadlocks in process state
$(Q\parallel[\Sigma]U_F(\# u))/u$, so it cannot produce the $\epass$-event;
this  means that the test fails and concludes the proof. 
\xbox
\end{proof}
